### 2014-12-29 external/notes/cropland-bias-notes.tex
Updated notes to reflect change from 20 to 25 km resolution, and land surface model bias example.

\begin{enumerate}
  \item Satellite-derived landcover data are a key ingredient for understanding many aspects of environmental, social, and economic change. Examples:
    \begin{itemize}
      \item Food production estimates (e.g. yield gaps) - anything by RamanFoley, Rachel's work, etc.
      \item Carbon stocks (Tier 1 vegetated carbon estimates) - Ruesch and Gibbs
      \item Habitat loss data.
      \item Attempts to estimate land-atmosphere interactions, hydrological models, etc. 
    \end{itemize}
  \item It is understood that there are errors inherent in these data, particularly global-scale products, and that there is great discrepancy between these. This is particularly true of developing regions that will experience the largest amount of land surface change this century, such as Africa.  
    \begin{itemize} 
      \item Cite to Fritz's papers (from geowiki). 
      \item Emphasize what they did so that we can differentiate--they look primarily at disagreements between products, and how much it affects country level estimates. Also focus on older generation datasets. 
    \end{itemize}
  \item What is less known is how much bias these landcover errors might introduce into research findings that depend on areas estimates drawn from landcover data. Such an assessment depends on having accurate, spatially extensive landcover data to assess large areas, which are typically not available. 
  \item In this study, we use a unique, high accuracy agricultural landcover dataset to assess the biases in the most current landcover datasets, looking at bias as a function of spatial scale and landcover fraction.  
  \item Our goal is to provide insight--for users of landcover products across a range of disciplines--into: 
    \begin{itemize}
      \item The scales at which area-based inferences drawn from global landcover products are appropriate
      \item How variations within landcover properties may influence bias
      \item Which datasets are most appropriate for making inferences
    \end{itemize}
\end{enumerate}

@@

%\vspace{0.5 in}
\begin{figure}[H]
\includegraphics[width = \linewidth]{../figures/bias_map.pdf}
\caption{The spatial distribution of biases in cropland cover estimates from four landcover products, calculated by subtracting each of their cropland percentage estimates from that of GeoTerraImage's cropland data layer. The biases were calculated for several different levels of, ranging from the original 1 km resolution up to 300 km (only 1, 20, 50, and 100 km results are shown here). Blue areas are those where the landcover products underestimated croplands, and red areas are where they overestimated. The GlobCover and MODIS results are drawn from biases calculated from the intermediate cropland fraction estimates per pixel.}
\end{figure}
%\vspace{-0.3in}

>>>>

\begin{enumerate}
  \item Satellite-derived landcover data are a key ingredient for understanding many aspects of environmental, social, and economic change. Examples:
    \begin{itemize}
      \item Food production estimates (e.g. yield gaps) - anything by RamanFoley, Rachel's work, etc.
      \item Carbon stocks (Tier 1 vegetated carbon estimates) - Ruesch and Gibbs
      \item Habitat loss data.
      \item Attempts to estimate land-atmosphere interactions, hydrological models, etc. [Here we will run land surface model with GTI cropland fraction providing the cropland surface, and we will then alter this using the bias rasters to see how this propagates into ET and runoff estimation errors, etc] 
    \end{itemize}
  \item It is understood that there are errors inherent in these data, particularly global-scale products, and that there is great discrepancy between these. This is particularly true of developing regions that will experience the largest amount of land surface change this century, such as Africa.  
    \begin{itemize} 
      \item Cite to Fritz's papers (from geowiki). 
      \item Emphasize what they did so that we can differentiate--they look primarily at disagreements between products, and how much it affects country level estimates. Also focus on older generation datasets. 
    \end{itemize}
  \item What is less known is how much bias these landcover errors might introduce into research findings that depend on areas estimates drawn from landcover data. Such an assessment depends on having accurate, spatially extensive landcover data to assess large areas, which are typically not available. 
  \item In this study, we use a unique, high accuracy agricultural landcover dataset to assess the biases in the most current landcover datasets, looking at bias as a function of spatial scale and landcover fraction.  
  \item Our goal is to provide insight--for users of landcover products across a range of disciplines--into: 
    \begin{itemize}
      \item The scales at which area-based inferences drawn from global landcover products are appropriate
      \item How variations within landcover properties may influence bias
      \item Which datasets are most appropriate for making inferences
    \end{itemize}
\end{enumerate}

@@

%\vspace{0.5 in}
\begin{figure}[H]
\includegraphics[width = \linewidth]{../figures/bias_map.pdf}
\caption{The spatial distribution of biases in cropland cover estimates from four landcover products, calculated by subtracting each of their cropland percentage estimates from that of GeoTerraImage's cropland data layer. The biases were calculated for several different levels of, ranging from the original 1 km resolution up to 300 km (only 1, 25, 50, and 100 km results are shown here). Blue areas are those where the landcover products underestimated croplands, and red areas are where they overestimated. The GlobCover and MODIS results are drawn from biases calculated from the intermediate cropland fraction estimates per pixel.}
\end{figure}
%\vspace{-0.3in}
